---
title: Final Project Code: Spotify Tracks
author: Mina Mahmoodzadeh
subtitle: MGSC 310
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = FALSE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```




```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('tidyverse')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Summary Plot 1

Here is a summary plot for our data set that is a bar chart that relates average popularity to the top 10 genres 

```{r}

# code for summary plot 1
library(ggplot2)
library(dplyr)

# loading in the spotify track dataset
file_path <- "spotifytracks.csv" 
dataset <- read.csv(file_path)

# calculating average popularity for each genre
genre_avg_popularity <- dataset %>%
  group_by(track_genre) %>%
  summarise(avg_popularity = mean(popularity, na.rm = TRUE)) %>%
  arrange(desc(avg_popularity)) %>%
  slice(1:10) # selecting the top 10 genres by average popularity

# create the bar chart
ggplot(genre_avg_popularity, aes(x = reorder(track_genre, avg_popularity), y = avg_popularity)) +
  geom_bar(stat = "identity", fill = "darkgreen", color = "black") +
  coord_flip() +
  labs(
    title = "Top 10 Genres by Average Popularity",
    x = "Genre",
    y = "Average Popularity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12)
  )

```

## Summary Plot 2 and 3

Here is our second and third summary plot which are boxplots that show the relationship between danceability and energy in relation to popularity. Popularity is split into three different sections (low, medium, and high popularity) for ease to read/analyze. 

```{r}

# code for summary plot 2 and 3

library(ggplot2)
library(dplyr)

# loading the dataset 
file_path <- "spotifytracks.csv"
dataset <- read.csv(file_path)

# Step 1: Create popularity ranges
dataset <- dataset %>%
  mutate(
    popularity_range = cut(popularity,
                           breaks = c(0, 30, 60, 100),
                           labels = c("Low", "Medium", "High"),
                           include.lowest = TRUE)
  )

# Step 2: Boxplot for Energy by Popularity Range
ggplot(dataset, aes(x = popularity_range, y = energy, fill = popularity_range)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Greens") +
  labs(
    title = "Energy Distribution by Popularity Range",
    x = "Popularity Range",
    y = "Energy"
  ) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

# Step 3: Boxplot for Danceability by Popularity Range
ggplot(dataset, aes(x = popularity_range, y = danceability, fill = popularity_range)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Greens") +
  labs(
    title = "Danceability Distribution by Popularity Range",
    x = "Popularity Range",
    y = "Danceability"
  ) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))


```


## Predictive Analysis Portion: Logistic Regression

 


```{r}

# code for logistic regression predictive analysis

library(ggplot2)
library(dplyr)
library(caret)

# loading in the dataset
file_path <- "spotifytracks.csv"
dataset <- read.csv(file_path)

# Defining the top 10% threshold
top_10_threshold <- quantile(dataset$popularity, 0.90)

# Create a binary variable: 1 for top 10%, 0 otherwise
dataset <- dataset %>%
  mutate(top_10 = ifelse(popularity >= top_10_threshold, 1, 0))

# Select predictor features and the target variable
features <- c("energy", "danceability", "acousticness", "tempo", "liveness", "instrumentalness")
target <- "top_10"

# Split the dataset into training (70%) and testing (30%) sets
set.seed(123) # For reproducibility
train_index <- createDataPartition(dataset$top_10, p = 0.7, list = FALSE)
train_data <- dataset[train_index, ]
test_data <- dataset[-train_index, ]

# Train a logistic regression model
logistic_model <- glm(top_10 ~ energy + danceability + acousticness + tempo + liveness + instrumentalness,
                      data = train_data,
                      family = binomial)

# Summary of the model
summary(logistic_model)

# Make predictions on the test set
predictions <- predict(logistic_model, newdata = test_data, type = "response")
test_data$predicted <- ifelse(predictions > 0.5, 1, 0)

# Confusion Matrix
confusion_matrix <- confusionMatrix(as.factor(test_data$predicted), as.factor(test_data$top_10))
print(confusion_matrix)

# ROC Curve and AUC
library(pROC)
roc_curve <- roc(test_data$top_10, predictions)
plot(roc_curve, main = "ROC Curve")
auc(roc_curve)

# Extract and plot feature importance
coefficients <- coef(summary(logistic_model))
feature_importance <- data.frame(Feature = rownames(coefficients),
                                 Coefficient = coefficients[, "Estimate"])
ggplot(feature_importance, aes(x = reorder(Feature, Coefficient), y = Coefficient)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  coord_flip() +
  labs(title = "Feature Importance (Logistic Regression)",
       x = "Features",
       y = "Coefficient")

```


2b) Response to part b. 


```{r}

# code for 2b

```


## Question 3

Here is only text for question 3. 

If you want to add a hypertext link [you can do so](http://www.chapman.edu)

- If you want to add bullets
  - You
  - Can 
  - Do 
  - It

If you want to add an image [follow this guide](https://stackoverflow.com/questions/25166624/insert-picture-table-in-r-markdown)



